Sediment Interception by Emergent Stems Across Varying Patch Densities and Flows

### 2021-01-08 aa workingdraft/draft.tex
moved preamble back to draft.tex

\input{preamble}

@@

To maximize the positive effect of managed vegetation on elevation gain in wetlands and estuaries threatened by sea level rise, one must consider effects of vegetation properties on both capture on plant surfaces and settling to the bottom of the channel. The quantity that best represents the capture efficiency of the vegetation patch as a whole is $a \eta^\prime$, where $a = I_cd_c$ is collector frontal-area density. For emergent stems modeled as cylinders, $a = \frac{4\phi_c}{\pi d_c}$ can be deduced geometrically ($\phi_c = I_c\frac{\pi}{4}d_c^2$). As a thought experiment: if $\phi_c$ is held constant, $a$ is inversely proportional to $d_c$, so smaller stems would result in increased frontal-area density. If $u$ is additionally held constant, Eq. (\ref{eq:fauria}) predicts a negative relationship between $d_c$ and ECE ($\eta^\prime \sim \Rey_c^{-1.14}R^{0.65} \sim d_c^{-1.14 - 0.65} \sim d_c^{-1.79}$). Since our results support the validity of Eq. (\ref{eq:fauria}), we conclude that minimizing $d_c$ is predicted to increase the total amount of sediment captured on plant surfaces in vegetation patches, at least in the case of emergent stems.

@@

For actual vegetation, $d_c$ may vary within a given vegetation patch, which may affect turbulence and thereby ECE. Because we only observed patches with uniform $d_c$, our study does not provide information on this additional complication. There is also evidence of a positive relationship between $d_c$ and collector density for actual aquatic plant species and communities \cite{mudd2010does}. Since there is convincing evidence that reduced collector density results in reduced total capture rate (including settling) \cite{Fauria_2015}, efforts to minimize $d_c$ should only go as far as is possible without also reducing collector density. Examples of plants with

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

>>>>

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{TBDXX}
\issuenum{TBDYY}
\articlenumber{TBDZZ}
\pubyear{2021}
\copyrightyear{2021}
\externaleditor{Academic Editor: Dr. Rafael Tinoco}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

\newcommand\Rey{\mathrm{Re}}
\newcommand\Pec{\mathrm{Pe}}

\usepackage{siunitx}
\DeclareSIUnit\year{yr}

\usepackage{textcomp}

\usepackage{threeparttable}
\usepackage{longtable}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Sediment Interception by Emergent Stems Across Varying Patch Densities and Flows}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-7970-841X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Jordan Wingenroth $^{1}$*\orcidA{}, Candace Yee $^{2}$, Justin Nghiem$^{1,3\dagger}$ and Laurel Larsen $^{1,2}$}

% Authors, for metadata in PDF
\AuthorNames{Jordan Wingenroth, Candace Yee, Justin Nghiem and Laurel Larsen}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Geography, University of California, Berkeley, CA

$^{2}$ \quad Department of Civil and Environmental Engineering, University of California, Berkeley, CA

$^{3}$ \quad Department of Statistics, University of California, Berkeley, CA}

% Contact information of the corresponding author
\corres{Correspondence: j.wingenroth@berkeley.edu}

% Current address and/or shared authorship
\firstnote{Current address: California Institute of Technology, Pasadena, CA} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Suspended sediment collected by vegetation in marshes and wetlands contributes to vertical accretion, which can buffer against rising sea levels and preserve these ecosystems. Effective capture efficiency (ECE), a parameter relating the rate of sediment capture by submerged vegetation surfaces to their effective diameters, is a key variable in determining the significance of direct interception in elevation-change models. The factors governing ECE are not yet thoroughly understood in transitionally turbulent flows. Using a recirculating flume containing an array with a variable count of artificial cylindrical collectors, we conducted a set of 12 experiments at 3 flow speeds and 3 collector densities (plus control), plus 4 additional experiments where biofilm was allowed to accumulate. We determined that ECE decreases with increasing collector Reynolds number (study range: 66 to 200; p < 0.05 for 2 of 3 treatments) and collector density (solid volume fraction: 0.22\% to 1.17\%; p < 0.05 for 2 of 3 treatments). Adding biofilm increased ECE in all cases, by a multiplicative factor ranging from 1.53 to 7.15 at different collector densities and biofilm growth durations. By combining our data with those of the most similar study, we also present a preliminary model quantitatively assessing the effect of collector density on ECE.}

% Keywords
\keyword{sediment transport; collector efficiency; submerged vegetation; transitional turbulence; biofilm; sedimentation}

@@

To maximize the positive effect of managed vegetation on elevation gain in wetlands and estuaries threatened by sea level rise, one must consider effects of vegetation properties on both capture on plant surfaces and enhanced settling of particles to the bottom of the channel. The quantity that best represents the capture efficiency of the vegetation patch as a whole is $a \eta^\prime$, where $a = I_cd_c$ is collector frontal-area density. For emergent stems modeled as cylinders, $a = \frac{4\phi_c}{\pi d_c}$ can be deduced geometrically ($\phi_c = I_c\frac{\pi}{4}d_c^2$). As a thought experiment: if $\phi_c$ is held constant, $a$ is inversely proportional to $d_c$, so smaller stems would result in increased frontal-area density. If $u$ is additionally held constant, Eq. (\ref{eq:fauria}) predicts a negative relationship between $d_c$ and ECE ($\eta^\prime \sim \Rey_c^{-1.14}R^{0.65} \sim d_c^{-1.14 - 0.65} \sim d_c^{-1.79}$). Since our results support the validity of Eq. (\ref{eq:fauria}), we conclude that minimizing $d_c$ is predicted to increase the total amount of sediment captured on plant surfaces in vegetation patches, at least in the case of emergent stems.

@@

For actual vegetation, $d_c$ may vary between different stems within a given vegetation patch, which may affect turbulence and thereby ECE. Because we only observed patches with uniform $d_c$, our study does not provide information on this additional complication. There is also evidence of a positive relationship between $d_c$ and collector density for actual aquatic plant species and communities \cite{mudd2010does}. Since there is convincing evidence that reduced collector density results in reduced total capture rate (including settling) \cite{Fauria_2015}, efforts to minimize $d_c$ should only go as far as is possible without also reducing collector density. Examples of plants that inhabit coastal estuaries and wetlands and that have relatively high $a$ values and relatively low $d_c$ include: xxx.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{Raw data and computer code for this study are available publicly on our \href{https://github.com/wingfactor/wingenroth-et-al-2021}{repository}.}
